hatchR: A toolset to predict hatch and emergence phenology in wild fishes

### 2025-03-06 inst/manuscript/paper.Rmd
add new figure 2

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse); library(hatchR); library(ggdist); library(ggridges)
```

@@

@@

```{r plot-check-1, echo = TRUE, fig.cap="Output of hatchR function `plot_check_temp()`, which is used as a visual data check on the raw `year_sim` data set. Users can set custom thresholds for minimum and maximum temperatures (dashed lines)."}
plot_check_temp(data = year_sim,
                dates = date,
                temperature = temp,
                temp_min = 0, # temp_min and max lines are 
                temp_max = 25) # user customizable
```

>>>>

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse); library(hatchR); library(ggdist); library(ggridges); library(patchwork)
```

@@

To demonstrate this approach, we use European Grayling (*Thymallus thymallus*) that were raised in five experimental average temperatures ranging from 2-14 °C [@quinn2018]. We fit a developmental model to those experimental data (Figure 2A), reparameterize that model into an effective value model, and predict effective values for daily average temperatures ranging from 2-14 °C (Figure 2B).

```{r, fig.cap= "Effective value model fit and application for European Grayling. Panel A shows developmental rates (days to hatch) at five differrent average rearing temperatures and the corresponding model fit (blue line). Panel B uses the effective value model derived from the blue line in panel A to predict the effective value of daily average temperatures ranging from 2 to 14 °C. The dashed line is a 0.01 increase in effective value for every 1 °C increase as a point of reference.",fig.height= 8, fig.width=4,  fig.align = 'center', echo =FALSE}

grayling <- matrix(NA, 5, 2) |> tibble::tibble()
colnames(grayling) <- c("temp", "days")

grayling$temp <- c(2,5,8,11,14)
grayling$days <- c(106,51,29,19,13)

grayling_mod <- hatchR::fit_model(temp = grayling$temp,
                                  days = grayling$days,
                                  species = "grayling",
                                  development_type = "hatch")

temps <- c(2:14)

ef_vals <- NULL
for (x in temps){
  ef <- eval(parse(text = grayling_mod$expression$ expression))
  ef_vals <- rbind(ef_vals, ef)
}

grayling_ef <- matrix(NA, 13, 2) |> tibble::tibble()
colnames(grayling_ef) <- c("temp", "ef")
grayling_ef$temp <- temps
grayling_ef$ef <- ef_vals[,1]

p1 <- grayling_mod$pred_plot +
  labs(x = "Developmental average temperature (°C)", y = "Days to hatch") +
  theme_classic()

data_1 <- tibble(t = c(0:14), e = seq(0, 0.14, by =0.01))

p2 <- grayling_ef |>
  ggplot() +
  geom_point(aes(x = temp, y = ef)) +
  geom_line(aes(x = temp, y = ef)) +
  geom_line(data = data_1, aes(x = t, y = e), linetype = "dashed") +
  #geom_abline(intercept = 0, slope = .01, linetype = "dashed") +
  labs(x = "Daily average temperature (°C)", y = "Effective value") +
  theme_classic()

p1/p2 + plot_annotation(tag_levels = 'A')

```

@@

```{r plot-check-1, echo = TRUE,  fig.cap="Output of hatchR function `plot_check_temp()`, which is used as a visual data check on the raw `year_sim` data set. Users can set custom thresholds for minimum and maximum temperatures (dashed lines).", fig.height= 4, fig.width=6.5}
plot_check_temp(data = year_sim,
                dates = date,
                temperature = temp,
                temp_min = 0, # temp_min and max lines are 
                temp_max = 25) # user customizable
```
